\input{labspre.tex}

\usepackage[compact]{titlesec}

\begin{document}
\MYTITLE{Laboratory Assignment Six: Automatically Characterizing Programs and Test Suites}
\MYHEADERS{Laboratory Assignment Six}{Due: November 6, 2013}

\section*{Introduction}

In the previous laboratory assignments, you have learned about the tools that we will use during the remainder of this
semester to specify, design, implement, test, and document Java programs.  You have also had several experiences with
working in progressively larger teams to complete the phases of the software life cycle, with a recent focus on the
elicitation of software requirements and the planning of software projects.  In this assignment, you and your team will
install, configure, and use software tools that automatically calculate design and implementation metrics that
characterize programs and test suites.  One week from now, you will present your findings from your analyses.  

\section*{Calculating Design Quality Metrics with JDepend}

After you have written a requirements document for your system and completed the description of the architecture, you
must create a design for your program.  Of course, it is important to evaluate the quality of your design.  If you
already have an implementation of a program and you want to evaluate the quality of its design, you can use a tool
called JDepend.  You can learn more about JDepend by visiting \url{http://clarkware.com/software/JDepend.html}. What are
the design quality metrics that JDepend calculates? What are the meanings of these metrics? How can you use these
metrics to better understand and modify the design of your Java programs? As one of the deliverables for this
assignment, your team should use \LaTeX\ to prepare a document that contains formal definitions and equations that
describe all of the design quality metrics. 

Once you and your team have investigated and discussed the features provided by JDepend, you should pick two separate
systems that have been previously implemented by members of your team. Now, copy the source code from these two projects
--- and any other files that your team deem to be relevant --- to the version control repository that you will use for
this project.  After downloading and installing JDepend, you should add rules to each project's Ant build system that
can run the JDepend analysis on the byte code. What does the output of JDepend tell you about the quality of each
project's design? Do the chosen programs have good designs?

\section*{Source Code Measurement with JavaNCSS}

As you are implementing your program, it is important to regularly calculate metrics that characterize the quality of
your implementation.  JavaNCSS is a software tool that can automatically scan your Java source code and report
information about the number of non-commented source code statements and the cyclomatic complexity.  You can learn more
about JavaNCSS by visiting \url{http://www.kclee.de/clemens/java/javancss/}. After reading the Web site for JavaNCSS,
you should search for papers in the ACM Digital Library, available at \url{http://dl.acm.org/}, that formally describe
the meaning of metrics calculated by JavaNCSS, such as cyclomatic complexity. As one of the deliverables for this
assignment, your team should use \LaTeX\ to write a document that contains formal definitions and equations that explain
all of the metrics calculated by JavaNCSS.  

Once all of the members of your team understand the metrics calculated by JavaNCSS, you should download and install
this tool.  As you did with JDepend, now you must add rules to each project's build system that can run JavaNCSS on the
source code. What does the output of JavaNCSS tell you about the quality of each project's implementation? Do your
chosen programs exhibit good implementation characteristics? How can you use the values of these metrics to better
understand and modify the implementation of your Java programs? How much of your system's source code is for the
program itself? How much of the source code is devoted to the test suite?

\section*{Summary of the Required Deliverables}

This assignment invites your team to submit one printed version of the following files:
\vspace*{-.1in}
\begin{enumerate}
	% \itemsep0em 
	\item A description of and justification for your team's chosen organization, roles, and tool support
	\item A document that clearly explains the meaning of JDepend's design quality metrics 
	\item An analysis of the values of design quality metrics for two previously implemented systems 
	\item A document that clearly explains the meaning of metrics calculated by JavaNCSS
	\item An analysis of the values of the implementation metrics for the two systems that you chose
	\item Any modified implementation artifacts (e.g., build system and source code) 
\end{enumerate}
\vspace*{-.1in}

You must also ensure that the instructor has read access to your Bitbucket repository that is named according to the
convention {\tt cs290F2013-lab6-team{\em k}}, with {\tt {\em k}} representing the number of your assigned team.  Your
repository should contain all of the deliverables that you produced during the completion of this assignment.  Please
see the instructor if you have any questions.

\end{document}
