\input{syllabuspre.tex}

\usepackage[compact]{titlesec}

\begin{document}
\MYTITLE{Syllabus}
\MYHEADERS{Syllabus}{}

\subsection*{Course Instructor}
Dr.\ Gregory M.\ Kapfhammer\\
\noindent Office Location: Alden Hall 108 \\
\noindent Office Phone: +1 814-332-2880 \\
\noindent Email: \url{gkapfham@allegheny.edu} \\
\noindent Web Site: \url{http://www.cs.allegheny.edu/sites/gkapfham/}

\subsection*{Instructor's Office Hours}

\begin{itemize}
	\itemsep 0em
	\item Monday: 1:00 pm -- 2:30 pm (30 minutes time slot)
	\item Tuesday: 2:30 pm -- 4:00 pm (15 minutes time slot)
	\item Wednesday: 4:30 pm -- 5:00 pm (15 minutes time slot)
	\item Thursday: 9:00 am -- 10:00 am (15 minutes time slot) {\em and} \\ \hspace*{.69in} 2:30 pm -- 4:00 pm (15 minutes time slot)
	\item Friday: 1:00 pm -- 2:30 pm (10 minutes time slot) {\em and} \\ \hspace*{.49in} 4:30 pm -- 5:00 pm (5 minutes time slot)
\end{itemize}

\noindent
To schedule a meeting with me during my office hours, please visit my Web site and click the ``Schedule'' link in the
top right-hand corner. Now, you can browse my office hours or schedule an appointment by clicking the correct link and
then reserving an open time slot. 

\subsection*{Course Meeting Schedule}

Lecture, Discussion, Presentations, and Group Work: Tuesday and Thursday 11:00 am -- 12:15 pm \\
Laboratory Session: Wednesday 2:30 pm -- 4:20 pm \\
Final Examination: Friday, December 13, 2013 at 9:00 am


\subsection*{Course Catalogue Description}

\begin{quote}

	A study of the principles and concepts used in the specification, design, implementation, testing, and maintenance
	of large software systems. Topics include requirements elicitation and analysis, formal specification, software
	architectures, object-oriented design, software measurement, software testing and analysis, and evolution of a
	program.  Students practice the principles of software development by participating as group members in the creation
	of a significant software application. One laboratory per week. Prerequisites: Computer Science 210 and 220 or
	permission of the instructor. \\ {\em Offered in alternate years}.

\end{quote}

\subsection*{Course Objectives}

The process of developing software involves the application of a number of interesting theories, tools, techniques, and
methodologies.  In this class we will explore the phases of the major software development life cycles and examine the
concepts, tools, challenges, and open questions associated with each phase.  Throughout the semester, we will
examine the interplay between the theory and practice of software development.  We will delve into the details of
software specification, design, implementation, testing, and maintenance through a discussion of articles and book
chapters from the software engineering and software testing literature.  Moreover, students will enhance their ability
to write about software in a clear and concise fashion.  Finally, students will gain practical software development
experience in laboratory sessions and a final software development project.

\subsection*{Performance Objectives}

At the completion of this class, a student should be aware of the fundamental challenges associated with software
development.  Furthermore, a student should be comfortable with a wide array of tools, techniques, concepts, and
methodologies that they can apply to the problem of developing large software systems.  However, a successful student
will emerge with more than an understanding of the tools (e.g., text editors, compilers, debuggers, integrated
development environments, and version control systems) that a software engineer uses.  A student also should have a
fundamental understanding of the major software life cycles and the activities that take place in each phase of these
life cycles.  Finally, a student should have a basic understanding of some of the current research and the open research
questions in the field of software engineering.  After completing this class, a student should be equipped for further
graduate study in the fields of computer science and software engineering.  The student should also be able to
participate in real-world software development projects by adeptly using modern software tools and working with a team of
developers.

\subsection*{Required Textbooks}

% Shari Lawrence Pfleeger and Joanne M. Atlee
%   Software Engineering: Theory and Practice (Fourth Edition)
%   ISBN-10: 0136061699
%   ISBN-13: 978-0136061694
%   Prentice Hall
%   Status: Required
%   25 copies

\noindent{\em Software Engineering: Theory and Practice}. Shari Lawrence Pfleeger and Joanne M. Atlee.
Fourth Edition, ISBN-10: 0136061699, ISBN-13: 978-0136061694, 792 pages, 2009. \\
(References to the textbook are abbreviated as ``SETP'' in the syllabus and on the Web site).

 % The Mythical Man-Month: Essays on Software Engineering, Anniversary
 %  Edition (2nd Edition) [Paperback]
 %  Author: Frederick P. Brooks
 %  Publisher: Addison-Wesley Professional; Anniversary edition (August 12, 1995)
 %  ISBN-10: 0201835959
 %  ISBN-13: 978-0201835953
 %  Status: Required
 %  25 copies

\noindent{\em The Mythical Man Month}. Frederick P.\ Brooks, Jr.\.
Second Edition, ISBN-10: 0201835959, ISBN-13: 978-0201835953, 336 pages, 1985. \\
(References to the textbook are abbreviated as ``MMM'' in the syllabus and on the Web site).

\noindent
Students who want to improve their technical writing skills may consult the following books.

\noindent{\em BUGS in Writing: A Guide to Debugging Your Prose} 
(Second Edition). Lyn Dupr\'e.  Addison-Wesley Professional.
ISBN-10: 020137921X and ISBN-13: 978-0201379211, 704 pages, 1998.
%References to the textbook are abbreviated as "BIW"

\noindent{\em Writing for Computer Science} (Second Edition).
Justin Zobel.  Springer
ISBN-10: 1852338024 and ISBN-13: 978-1852338022, 270 pages, 2004.
%References to the textbook are abbreviated as "WFCS"

Along with reading the required books, you will be asked to study many
additional papers from a wide variety of conference proceedings and
journals.

\subsection*{Grading Policy}

% The grade that a student receives in this class will be based on class
% participation, grades on two interim exams, the grade on the final
% exam, the grade for all of the laboratory and homework assignments,
% and the grade for the final project.  The grade breakdown will follow
% the form below.  All percentages are approximate and it is possible
% for the assigned percentages to change if a need to do so presents
% itself.

The grade that a student receives in this class will be based on the
following categories. All percentages are approximate and it is possible
for the assigned percentages to change during the academic semester if a
need to do so presents itself. 

\begin{center}
\begin{tabular}{ll}
Class Participation and Instructor Meetings & 5\% \\
First Examination & 15\% \\
Second Examination & 15\% \\
Final Examination & 20\% \\
Laboratory and Homework Assignments & 30\% \\
Final Project & 15\%
\end{tabular}
\end{center}

Each of the above grading categories has the following definition:

\begin{itemize}

	\item {\em Class Participation and Instructor Meetings}: All students are required to actively participate during
		all of the class sessions. Your participation will take forms such as answering questions about the required
		reading assignments, asking constructive questions of your group members, giving presentations, and leading a
		brainstorming session. All students are required to meet with the course instructor for a total of sixty minutes
		during the Fall 2013 semester.  These meetings must be scheduled through the instructor's reservation system and
		documented on a meeting record that you submit on the day of the final examination. A student will receive an interim 
		and final grade for this category.

	\item {\em First and Second Examinations}: The first and second interim examinations will cover all of the material
		in the associated module(s).  While the second examination is not cummulative, it will assume that a student has a
		basic understanding of the material that was the focus of the first examination.  The date for the first and
		second examinations will be announced at least one week in advanced of the scheduled date.  Unless prior
		arrangements are made with the course instructor, all students will be expected to take the examination on the
		scheduled date and complete the test in the stated period of time.

	\item {\em Final Examination}: The final examination is a three-hour cummulative test with a time slot 

	\item {\em Laboratory Assignments}: will require students to explore the concepts, tools, and techniques that are
		associated with different phases of the software development lifecycle.  All laboratory assignments will require
		the use of the provided tools to design, implement, test, and maintain programs that solve important problems.
		Each laboratory assignment will normally be provided on the day before our scheduled laboratory session.  The
		Instructor expects students to read the laboratory assignment before attending the session to prepare themselves
		appropriately.  Unless specified otherwise, each laboratory assignment will be due at the beginning of the next
		laboratory session.  Many of the laboratory assignments in this course will require the students to give a
		presentation and a demonstration of the software that they implemented and tested.  

\end{itemize}

\subsection*{Assignment Submission Policy}

All assignments will have a given due date. The printed version of the
assignment is to be turned in at the beginning of the class on that due
date. Late assignments will be accepted for up to one week past the
assigned due date with a 10\% penalty. All late assignments must be
submitted at the beginning of the class that is scheduled one week after
the given due date. Unless special arrangements are made with the course
instructor, no assignments will be accepted after the late deadline.

\subsection*{Class Attendance Policy}

It is mandatory for all students to attend class. If you will not be
able to attend a class session, then please see the course instructors at
least one week in advance in order to describe your situation. Students
who miss more than five unexcused classes will have their final grade in
the course reduced by one letter grade. Students who miss more than ten
unexcused classes will automatically fail the course.

\subsection*{Laboratory Attendance Policy}

In order to acquired the proper skills in technical writing, critical
reading, and the presentation of technical material, it is essential
for students to have hands-on experience in a laboratory. Therefore,
it is mandatory for all students to attend the laboratory sessions. If
you will not be able to attend a laboratory, then please see the
course instructors at least one week in advance in order to explain
your situation. Students who miss more than two unexcused laboratories
will have their final grade in the course reduced by one letter grade.
Students who miss more than four unexcused laboratories will
automatically fail the course.
\end{document}
